Supplementary Information to "Improving Models to Predict Holocellulose and Klason Lignin Contents for Peat Soil Organic Matter with Mid Infrared Spectra

### 2022-10-26 analysis/paper/001-paper-main.Rmd
Adapt color schemes for color blind persons.

  We like to thank @Hodgkins.2018 and @DelaCruz.2016 for provision of data and code to reproduce and modify their analyses, and @Allaire.2020 for provision of an Rmarkdown template. We thank Suzanne Hodgkins and Jeff Chanton for constructive discussions which helped to improve a first version of this manuscript. This study was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) grant no. KN 929/23-1 to Klaus-Holger Knorr and grant no. PE 1632/18-1 to Edzer Pebesma. We acknowledge support from the Open Access Publication Fund of the University of Münster.

@@

# table print options
options(knitr.table.format = "latex")
```

>>>>

  We like to thank @Hodgkins.2018 and @DelaCruz.2016 for provision of data and code to reproduce and modify their analyses, and @Allaire.2020 for provision of an Rmarkdown template. We thank Suzanne Hodgkins and Jeff Chanton for constructive discussions and Stephen Chapman and an anonymous reviewer for comments which helped to improve previous versions of this manuscript. This study was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) grant no. KN 929/23-1 to Klaus-Holger Knorr and grant no. PE 1632/18-1 to Edzer Pebesma. We acknowledge support from the Open Access Publication Fund of the University of Münster.

@@

# table print options
options(knitr.table.format = "latex")

# color theme (source: http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette)
palette_cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

### 2022-10-26 analysis/paper/003-paper-m-gaussian-beta.Rmd
Adapt color schemes for color blind persons.

p1 <-
  d_peat_pred %>%
  dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse()

p2 <-
  d_peat_pred %>%
  dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse()

@@

```{r m-gaussian-beta-6}
# plot
p1 <-
  d_range_pred %>%
  dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(x = fit_gaussian, y = fit, colour = model, fill = model)) +
  geom_hline(yintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_vline(xintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, colour = NA) +
  geom_path() +
  geom_point(data = d_peat_pred %>% dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit)) +
  geom_point(data = d_veg_pred %>% dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 4) +
  geom_point(data = d_range_pred_limits %>% dplyr::filter(variable_y == "Holocellulose"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 21, fill = "white", colour = "black") +
  labs(title = "Holocellulose", 
       x = expression(Predicted~values~"(Gaussian model)"),
       y = expression(Predicted~values)) +
  coord_cartesian(ylim = c(0, 1), xlim = c(0, 1)) +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100)

p2 <-
  d_range_pred %>%
  dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(x = fit_gaussian, y = fit, colour = model, fill = model)) +
  geom_hline(yintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_vline(xintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, colour = NA) +
  geom_path() +
  geom_point(data = d_peat_pred %>% dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit)) +
  geom_point(data = d_veg_pred %>% dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 4) +
  geom_point(data = d_range_pred_limits %>% dplyr::filter(variable_y == "Klason lignin"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 21, fill = "white", colour = "black") +
  labs(title = "Klason lignin", 
       x = expression(Predicted~values~"(Gaussian model)"),
       y = expression(Predicted~values)) +
  coord_cartesian(ylim = c(0, 1), xlim = c(0, 1)) +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100)

>>>>

p1 <-
  d_peat_pred %>%
  dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
  dplyr::mutate(
    model =
      dplyr::case_when(
        model == "Gaussian, Bayesian" ~ "Gaussian",
        TRUE ~ model
      )
  ) %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse() +
  guides(color = guide_legend(title = "Bayesian model"), fill = guide_legend(title = "Bayesian model")) +
  scale_color_manual(values = palette_cb[-1])

p2 <-
  d_peat_pred %>%
  dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
  dplyr::mutate(
    model =
      dplyr::case_when(
        model == "Gaussian, Bayesian" ~ "Gaussian",
        TRUE ~ model
      )
  ) %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse() +
  guides(color = guide_legend(title = "Bayesian model"), fill = guide_legend(title = "Bayesian model")) +
  scale_color_manual(values = palette_cb[-1])

@@

```{r m-gaussian-beta-6}
# plot
p1 <-
  d_range_pred %>%
  dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
  dplyr::mutate(
    model =
      dplyr::case_when(
        model == "Gaussian, Bayesian" ~ "Gaussian",
        TRUE ~ model
      )
  ) %>%
  ggplot(aes(x = fit_gaussian, y = fit, colour = model, fill = model)) +
  geom_hline(yintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_vline(xintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, colour = NA) +
  geom_path() +
  geom_point(
    data = 
      d_peat_pred %>% 
      dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
      dplyr::mutate(
        model =
          dplyr::case_when(
            model == "Gaussian, Bayesian" ~ "Gaussian",
            TRUE ~ model
          )
      ), 
    mapping = aes(x = fit_gaussian, y = fit)
  ) +
  geom_point(
    data = 
      d_veg_pred %>% 
      dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
      dplyr::mutate(
        model =
          dplyr::case_when(
            model == "Gaussian, Bayesian" ~ "Gaussian",
            TRUE ~ model
          )
      ), 
    mapping = aes(x = fit_gaussian, y = fit), shape = 4
  ) +
  geom_point(data = d_range_pred_limits %>% dplyr::filter(variable_y == "Holocellulose"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 21, fill = "white", colour = "black") +
  labs(title = "Holocellulose", 
       x = expression(Predicted~values~"(Gaussian model)"),
       y = expression(Predicted~values)) +
  coord_cartesian(ylim = c(0, 1), xlim = c(0, 1)) +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100) +
  guides(color = guide_legend(title = "Bayesian model"), fill = guide_legend(title = "Bayesian model")) + 
    scale_color_manual(values = palette_cb[-1]) +
    scale_fill_manual(values = palette_cb[-1])

p2 <-
  d_range_pred %>%
  dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
  dplyr::mutate(
    model =
      dplyr::case_when(
        model == "Gaussian, Bayesian" ~ "Gaussian",
        TRUE ~ model
      )
  ) %>%
  ggplot(aes(x = fit_gaussian, y = fit, colour = model, fill = model)) +
  geom_hline(yintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_vline(xintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, colour = NA) +
  geom_path() +
  geom_point(
    data = 
      d_peat_pred %>% 
      dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
      dplyr::mutate(
        model =
          dplyr::case_when(
            model == "Gaussian, Bayesian" ~ "Gaussian",
            TRUE ~ model
          )
      ), 
    mapping = aes(x = fit_gaussian, y = fit)
  ) +
  geom_point(
    data = 
      d_veg_pred %>% 
      dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
      dplyr::mutate(
        model =
          dplyr::case_when(
            model == "Gaussian, Bayesian" ~ "Gaussian",
            TRUE ~ model
          )
      ), 
    mapping = aes(x = fit_gaussian, y = fit), shape = 4
  ) +
  geom_point(data = d_range_pred_limits %>% dplyr::filter(variable_y == "Klason lignin"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 21, fill = "white", colour = "black") +
  labs(title = "Klason lignin", 
       x = expression(Predicted~values~"(Gaussian model)"),
       y = expression(Predicted~values)) +
  coord_cartesian(ylim = c(0, 1), xlim = c(0, 1)) +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100) +
  guides(color = guide_legend(title = "Bayesian model"), fill = guide_legend(title = "Bayesian model")) + 
    scale_color_manual(values = palette_cb[-1]) +
    scale_fill_manual(values = palette_cb[-1])

### 2022-10-26 analysis/paper/004-paper-m-reduce-underfitting.Rmd
Adapt color schemes for color blind persons.

```{r m-reduce-underfitting-9}
## plot depth profiles
p1 <-
  d_peat_pred_best %>%
  dplyr::filter(variable_y == "Holocellulose") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse()

p2 <-
  d_peat_pred_best %>%
  dplyr::filter(variable_y == "Klason lignin") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse()

>>>>

```{r m-reduce-underfitting-9}
## plot depth profiles
p1 <-
  d_peat_pred_best %>%
  dplyr::filter(variable_y == "Holocellulose") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse() +
  scale_color_manual(values = palette_cb[-1]) +
  scale_fill_manual(values = palette_cb[-1])

p2 <-
  d_peat_pred_best %>%
  dplyr::filter(variable_y == "Klason lignin") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse() +
  scale_color_manual(values = palette_cb[-1]) +
  scale_fill_manual(values = palette_cb[-1])

### 2022-10-26 analysis/paper/005-paper-m-minerals.Rmd
Adapt color schemes for color blind persons.

# plot
p_y_yhat_minerals <-
  d_minerals_pred %>%
  ggplot(aes(x = fit * 100, y = hol * 100, colour = mineral)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey") +
  geom_point() +
  coord_fixed() +
  facet_wrap( ~ model, nrow = 1) +
  labs(x = "Fitted",
       y = "Measured") +
  guides(colour = guide_legend(title = "Mineral rich samples")) +
  theme(legend.position = "bottom")
```

>>>>

# plot
p_y_yhat_minerals <-
  d_minerals_pred %>%
  ggplot(aes(x = fit * 100, y = hol * 100, colour = mineral)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey") +
  geom_point() +
  coord_fixed() +
  facet_wrap( ~ model, nrow = 1) +
  labs(x = "Fitted",
       y = "Measured") +
  guides(colour = guide_legend(title = "Mineral rich samples")) +
  theme(legend.position = "bottom") +
  scale_color_manual(values = palette_cb[-1])
```

### 2022-10-26 analysis/paper/007-paper-m-prediction-differences.Rmd
Adapt color schemes for color blind persons.

p1 <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = hol_hat * 100, 
             y = hol * 100)) +
  geom_point(aes(colour = sample_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Sample type")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Holocellulose") +
  coord_fixed() +
  scale_colour_manual(values = scales::hue_pal()(length(levels(d_pred_diff$sample_type))),
                    labels = levels(d_pred_diff$sample_type), 
                    drop = FALSE)

@@

p2 <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = kl_hat * 100, y = kl * 100)) +
  geom_point(aes(colour = sample_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Sample type")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Klason lignin") +
  coord_fixed() +
  scale_colour_manual(values = scales::hue_pal()(length(levels(d_pred_diff$sample_type))),
                    labels = levels(d_pred_diff$sample_type), 
                    drop = FALSE)

>>>>

p1 <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = hol_hat * 100, 
             y = hol * 100)) +
  geom_point(aes(colour = sample_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Sample type")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Holocellulose") +
  coord_fixed() +
  scale_color_manual(
    values = palette_cb,
    labels = levels(d_pred_diff$sample_type), 
    drop = FALSE
  ) +
  scale_fill_manual(
    values = palette_cb,
    labels = levels(d_pred_diff$sample_type), 
    drop = FALSE
  )

@@

p2 <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = kl_hat * 100, y = kl * 100)) +
  geom_point(aes(colour = sample_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Sample type")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Klason lignin") +
  coord_fixed() +
  scale_color_manual(
    values = palette_cb,
    labels = levels(d_pred_diff$sample_type), 
    drop = FALSE
  ) +
  scale_fill_manual(
    values = palette_cb,
    labels = levels(d_pred_diff$sample_type), 
    drop = FALSE
  )

### 2022-10-26 analysis/paper/008-paper-supplementary.Rmd
Adapt color schemes for color blind persons.

---
title: 'Supplementary Information to "Improving Models to Predict Holocellulose and Klason Lignin Contents for Peat Soil Organic Matter with Mid Infrared Spectra"'
author: "Henning Teickner, Klaus-Holger Knorr"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
header-includes:
   - \usepackage{float}
   - \renewcommand\thesection{S\arabic{section}}
---

@@

p2 <-
  d %>%
  ggplot(aes(y = kl * 100, x = hol * 100, colour = sample_type)) +
  geom_point() +
  labs(y = "Klason lignin [mass-%]", x = "Holocellulose [mass-%]") +
  geom_smooth(formula = y ~ x, method = "lm", se = FALSE) +
  guides(colour = guide_legend(title = "Sample type", ncol = 3, title.position="top")) +
  theme(legend.position = "bottom")
```

>>>>

---
title: 'Supplementary Information to "Improving Models to Predict Holocellulose and Klason Lignin Contents for Peat Soil Organic Matter with Mid Infrared Spectra"'
author: "Henning Teickner, Klaus-Holger Knorr"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
header-includes:
   - \usepackage{float}
   - \renewcommand\thesection{S\arabic{section}}
   - \renewcommand{\thefigure}{S\arabic{figure}}
---

@@

p2 <-
  d %>%
  ggplot(aes(y = kl * 100, x = hol * 100, colour = sample_type)) +
  geom_point() +
  labs(y = "Klason lignin [mass-%]", x = "Holocellulose [mass-%]") +
  geom_smooth(formula = y ~ x, method = "lm", se = FALSE) +
  guides(colour = guide_legend(title = "Sample type", ncol = 3, title.position="top")) +
  theme(legend.position = "bottom") +
  scale_color_manual(values = palette_cb)
```
